---
title: "Functional Analysis"
author: "Jannika"
date: "7/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Functional Analysis

```{r libraries}
if(!require("pacman")) install.packages("pacman")
pacman::p_load(data.table, dplyr, tidyverse, tibble, raster, sp, stars, sf, fossil, FD, multirich, matrixStats, ggplot2, ggspatial, myplot)
#input functions - file needs to be saved in the same folder as this Rmd
source('functional-analysis-functions.R')
```


#1. Assign Ecoregions

```{r ecoregions}
#load species data - output from species-distributions.Rmd
whales_raster <- readRDS('pres_rasterstack.Rds')

#make empty list for rasters
whales_shapefiles <- list()
#extract each raster and extend or crop extent accordingly
#this takes some time...once done save and load rather than redoing this bit every time...
for(i in 1:length(whales_raster@layers)){
  whales_shapefiles[[i]] <- sf::as_Spatial(sf::st_as_sf(stars::st_as_stars(whales_raster[[i]]),
                                                 as_points = FALSE, merge = TRUE)) 
}

saveRDS(whales_shapefiles, file = 'whales_shapefiles_list.Rds')
#whales_shapefiles <- readRDS('whales_shapefiles_list.Rds')

#Marine Ecoregions of the World (MEOW) obtained from:
#www.worldwildlife.org/publications/marine-ecoregions-of-the-world-a-bioregionalization-of-coastal-and-shelf-areas
MEOW <- shapefile('~/Sussex/thesis/MarineRegions/meow_ecos.shp')

#convert sptialpolygondataframe into sf object
MEOW <- st_as_sf(MEOW)
#create empty list and dataframe to fill with loop
MEOW_list <- list()
MEOW_DF <- data.frame()

sf::sf_use_s2(FALSE)

for(i in 1:length(whales_shapefiles)) {
  #make temporary sf object with first whale
  sf_temp <- st_as_sf(whales_shapefiles[[i]])
  #join ecoregion and first whale sf objects
  MEOW_temp <- st_join(MEOW, sf_temp)
  #add to list and convert to dataframe
  MEOW_list[[i]] <- MEOW_temp
  #select columm of ecoregion name and species
  MEOW_DF_temp <- as.data.frame(MEOW_temp[c(2,10)])
  row.names(MEOW_DF_temp) <- NULL
  MEOW_DF_temp$geometry <- NULL
  species <- colnames(MEOW_DF_temp)[2]
  colnames(MEOW_DF_temp)[2] <- "Species"
  MEOW_DF_temp <- gather(MEOW_DF_temp, Species, Measurment, Species)
  MEOW_DF_temp$Species <- species
  MEOW_DF <- rbind(MEOW_DF, MEOW_DF_temp)
  print(i) #print progress in console
}
#select only presence 
MEOW_DF <- MEOW_DF %>% filter(Measurment == 1) %>% unique()

write.csv(MEOW_DF, 'FctTradeoffs/species_MEOW.csv')
#MEOW_DF <- read.csv('FctTradeoffs/species_MEOW.csv')

#calculate species richness per ecoregion
eco_SR <- MEOW_DF %>% group_by(ECOREGION) %>% summarise(SR = sum(Measurment))

MEOW_with_SR <- merge(MEOW, eco_SR, by = "ECOREGION", all.x = TRUE)
saveRDS(MEOW_with_SR, 'FctTradeoffs/MEOW_with_SR.Rds')

#ggplot(data = MEOW_with_SR, aes(fill = SR)) +
#  geom_sf() +
#  scale_fill_viridis_b()

```

#2. Input ecoregion and traits data

```{r pressure}
# load ecoregion data

eco <- read.csv('FctTradeoffs/species_MEOW.csv')

eco <- eco %>% 
  select(c(ECOREGION, Species)) %>% 
  rename(eco = ECOREGION,
         species = Species)

eco$species <- sub('_',' ',eco$species)

saveRDS(eco, 'FctTradeoffs/eco_MEOW.Rds')
#eco <- readRDS('FctTradeoffs/eco_MEOW.Rds')

#load trait data
tr_mi <- readRDS('Traits/tr_mi.rds')
#check normality of continous traits
tr3 <- tr_mi[[3]] %>% 
  select(-species)
hist(tr3$YearsInterBirth) #continuous
hist(log10(tr3$YearsInterBirth)) #meh
hist(sqrt(tr3$YearsInterBirth)) #use this

hist(tr3$gestation_d) #continuous
hist(log10(tr3$gestation_d)) #perhaps a little better ?
lines(density(log10(tr3$gestation_d)), col = 2)
hist(sqrt(tr3$gestation_d)) #best
lines(density(sqrt(tr3$gestation_d)), col = 2)

hist(tr3$meanBodyMass) #continuous
hist(log10(tr3$meanBodyMass))
lines(density(log10(tr3$meanBodyMass)), col = 2)

#randomly select one of the 25 imputated datasets
td_mi <- tr_mi[[3]] %>% 
  #remove traits so I am only left with the 5 I want!
  dplyr::select(-YearsInterBirth, -breeding_site, -vertical, -DielActivity) %>% 
  dplyr::mutate_at(vars(meanBodyMass), log10) %>% 
  # log10-transform specified traits (body mass, litter size, generation length)
  dplyr::mutate(gestation = sqrt(gestation_d))
# square-root transform habitat breadth

#### Check loaded data ####

# species not found in any ecoregion
no_eco <- setdiff(td_mi$species, eco$species)
# Sousa chinensis - no aquamap for it thus can't assign ecoregion

# remove missing ecoregion species from trait data
td_mi <- td_mi %>% 
  filter(!species %in% no_eco)

# test if the site and trait data sets contain the same number of species? 
isTRUE(nrow(td_mi) == nrow(distinct(eco, species))) 
# should be TRUE

saveRDS(td_mi, 'FctTradeoffs/td_mi-5.Rds')
```

#3. 

```{r site data}
# create speciesxsite matrix
eco_mi_m <- t(create.matrix(as.data.frame(eco), tax.name = "species", locality = "eco"))

# add species pool
eco_mi_m_pool <- rbind(matrix(1, nrow = 1, ncol = ncol(eco_mi_m), 
                              dimnames = list(c("pool"), c(colnames(eco_mi_m)))), eco_mi_m)

# add pool as a site with every species present

# number of ecoregions assessed
e_mi <- dim(eco_mi_m)[1]
e_mi_pool <- e_mi+1
```

Z-transform trait data and bin continuous traits

```{r trait data transformation}
# function to z-transform data
scale_z <- function(x){
  (x - mean(x, na.rm=TRUE))/sd(x, na.rm=TRUE)
}

## td_mi_z ##

# z-transform trait data
td_mi_z <- td_mi %>%
  mutate_at(vars(-species), scale_z) %>% 
  # scale (zero mean and unit variance) all traits
  as.data.frame %>% 
  # convert to true dataframe ato prevent 'Warning message: Setting row names on a tibble is deprecated.'
  column_to_rownames(var = "species")
  # add species names to rownames (needed for gd function)

saveRDS(td_mi_z, 'FctTradeoffs/td_mi_z-5traits.Rds')
#td_mi_z <- readRDS('FctTradeoffs/td_mi_z.Rds')

# combine td_mi and eco_rm data
ts_mi <- list(td_mi_z, eco_mi_m_pool); names(ts_mi) <- c("td_mi","eco_mi")

# calculate gower species x species distance matrix based on traits
gd_mi <- gowdis(ts_mi$td_mi)

# function to bin data using sturges algorithm
cut_bin_stur <- function(x) {
  h <- hist(x, breaks = "Sturges")
  as.numeric(cut(x, breaks = h$breaks))
}

# bin data using sturges algorithm as the number of bins
tb_stur_mi <- td_mi %>% 
  mutate_at(vars(meanBodyMass, gestation_d), funs(cut_bin_stur)) %>% 
  # bin data
  as.data.frame %>% 
  # convert to true dataframe to prevent 'Warning message: Setting row names on a tibble is deprecated.'
  column_to_rownames(var = "species") 
  # convert species names to rownames

saveRDS(tb_stur_mi, 'FctTradeoffs/tb_stur_mi-5traits.Rds')
```

#4. Functional dispersion

```{r}
# calculate functional dispersion
fd_mi <- fdisp(gd_mi, ts_mi$eco_mi)

# modify functional dispersion output
fd_mi <- as.data.frame(fd_mi$FDis) %>%
  # extract data and convert to dataframe
  rownames_to_column("eco") %>% 
  # convert ecoregions to explicit column
  dplyr::rename(fdisp = `fd_mi$FDis`)
  # rename FDis column

saveRDS(fd_mi, "FctTradeoffs/fd_mi-5traits.rds")
```

#5. Functional redundancy

```{r fred}
# run fr_utc function
fr_mi <- fun_met(tdf = tb_stur_mi, em = eco_mi_m_pool, fd = fd_mi, e = e_mi_pool, name = "mi")

saveRDS(fr_mi, 'FctTradeoffs/MEOW_fr_mi-5traits.Rds')

mean(fr_mi$spp/fr_mi$utc) #1.046 for 5 traits
```

```{r}
colnames(fr_mi)[1] <- 'ECOREGION'

MEOW_with_SR <- readRDS('FctTradeoffs/MEOW_with_SR.Rds')

FR_MEOW_5 <- merge(MEOW_with_SR, fr_mi, by = "ECOREGION", all.x = TRUE)

saveRDS(FR_MEOW_5, 'FctTradeoffs/MEOW_FRed-5traits.Rds')
#FR_MEOW_5 <- readRDS('FctTradeoffs/MEOW_FRed-5traits.Rds')
```

##Null models
#1. upload all relevant data
```{r}
#upload data from functional-tradeoffs.R script
eco <- readRDS('FctTradeoffs/eco_MEOW.Rds')
#traits data from a single randomly selected imputation (3)
td_mi <- readRDS('FctTradeoffs/td_mi_5traits.Rds')
#z-transformed traits data
td_mi_z <- readRDS('FctTradeoffs/td_mi_z-5traits.Rds')
#binned traits data
tb_stur_mi <- readRDS('FctTradeoffs/tb_stur_mi-5traits.Rds')
```

#2. run null model
```{r random sample}
#1. get randomized species sample per ecoregion
# dataframe of global species trait data
spp <- td_mi %>%
  # scale (zero mean and unit variance) all traits
  mutate_at(vars(-species), scale_z) %>% 
  as.data.frame 

# dataframe of species richness per ecoregion
eco_spp <- table(eco$eco) %>% 
  # species richness per ecoregion
  as.data.frame %>% 
  # convert table to dataframe
  dplyr::rename(eco = Var1, spp = Freq)
# rename columns

# number of permutations to run
npermute <- 999

spp_eco <- replicate(
  npermute,
  eco_spp %>%
    dplyr::group_by(eco) %>%
    # nest columns within dataframe
    tidyr::nest() %>%
    # sample number of species equivalent to species richness
    dplyr::mutate(spp_list = purrr::map(data, ~ dplyr::sample_n(spp, .$spp))) %>%
    # unnest - turn data back into dataframe
    tidyr::unnest(cols = c(data, spp_list)),
  simplify = FALSE
)

#2. use existing code to calculate functional redundancy and functional dispersion for each randomized community
null <- list()

for(i in 1:length(spp_eco)){
eco_mi_m_spp <- t(create.matrix(as.data.frame(spp_eco[[i]]), tax.name = 'species', locality = 'eco'))

eco_mi_m_pool_spp <- rbind(matrix(1, nrow = 1, ncol = ncol(eco_mi_m_spp), 
                                  dimnames = list(c("pool"), c(colnames(eco_mi_m_spp)))), eco_mi_m_spp)

e_mi_spp <- dim(eco_mi_m_spp)[1]

e_mi_pool_spp <- e_mi_spp+1

ts_mi_spp <- list(td_mi_z, eco_mi_m_pool_spp); names(ts_mi_spp) <- c("td_mi","eco_mi")

gd_mi_spp <- gowdis(ts_mi_spp$td_mi)

fd_mi_spp <- fdisp(gd_mi_spp, ts_mi_spp$eco_mi)

# modify functional dispersion output
fd_mi_spp <- as.data.frame(fd_mi_spp$FDis) %>%
  # extract data and convert to dataframe
  rownames_to_column("eco") %>% 
  # convert ecoregions to explicit column
  dplyr::rename(fdisp = `fd_mi_spp$FDis`)

fr_mi_spp <- fun_met(tdf = tb_stur_mi, em = eco_mi_m_pool_spp, fd = fd_mi_spp, e = e_mi_pool_spp, name = "mi")

null[[i]] <- fr_mi_spp

}
#3. summarise the null model runs with the mean and standard deviation 

null_data <- rbindlist(null)
saveRDS(null_data, '~/R/thesis/FctTradeoffs/null_data_unlisted.Rds')
#null_data <- readRDS('~/R/thesis/FctTradeoffs/null_data_unlisted.Rds')
#unlist the dataframes into one giant dataframe
null_df <- null %>% 
  reduce(inner_join, by = 'eco')

#subset the dataframe for fdisp and calculate the mean fdisp per ecoregion
null_fdisp <- null_df %>% 
  select('eco', starts_with('fdisp')) %>% 
  mutate(fdisp_null = rowMeans(.[c(-1)])) %>% 
  mutate(fdisp_sd = rowSds(as.matrix(.[,c(-1)]))) %>% 
  select('eco', 'fdisp_null', 'fdisp_sd')

#subset the dataframe for fred and calculate the mean fred per ecoregion
null_fred <- null_df %>% 
  select('eco', starts_with('fred')) %>% 
  mutate(fred_null = rowMeans(.[c(-1)])) %>% 
  mutate(fred_sd = rowSds(as.matrix(.[,c(-1)]))) %>% 
  select('eco', 'fred_null', 'fred_sd')

#merge null model values into a single dataframe
Fnull <-  merge(null_fred, null_fdisp, by = 'eco', all.x = TRUE, all.y = TRUE)
saveRDS(Fnull, 'FctTradeoffs/Fnull_MEOW-5traits_df.Rds')
```

#3. calculate the Standardized Effect Size (SES)
SES = [Fobs - mean(Fnull)]/SD(Fnull)

```{r SES}
# functions for SES per ecoregion
SES_fred <- function(x){
  (x - null_fred$fred_null)/null_fred$fred_sd
}

SES_fdisp <- function(x){
  (x - null_fdisp$fdisp_null)/null_fdisp$fdisp_sd
}

#input observed funct metrics
Fobs <- readRDS('FctTradeoffs/MEOW_fr_mi-5traits.Rds')

#SES per ecoregion
SES <- Fobs %>% 
  mutate_at(vars(fred), SES_fred) %>% 
  mutate_at(vars(fdisp), SES_fdisp) %>% 
  select('eco', 'fred', 'fdisp') %>% 
  rename(fred_ses = fred,
         fdisp_ses = fdisp)

#add SES and Fobs to Fnull dataframe
Fct_df1 <- merge(Fnull, SES, by = 'eco', all.x = TRUE, all.y = TRUE)
Fct_df <- merge(Fobs, Fct_df1, by = 'eco', all.x = TRUE, all.y = TRUE)
Fct_df <- Fct_df %>% 
  rename(fred_obs = fred,
         fdisp_obs = fdisp)

saveRDS(Fct_df, 'FctTradeoffs/Fct_df_obs_null_SES-5traits.Rds')
#Fct_df <- readRDS('~/R/thesis/FctTradeoffs/Fct_df_obs_null_SES-5traits.Rds')

#in order to plot it merge with the MEOW sf file to get geometry values
#rename eco column to match MEOW df
colnames(SES)[1] <- 'ECOREGION'
#upload MEOW df
MEOW_with_SR <- readRDS('FctTradeoffs/MEOW_with_SR.Rds')

SES_MEOW <- merge(MEOW_with_SR, SES, by = "ECOREGION", all.x = TRUE)

saveRDS(SES_MEOW, 'FctTradeoffs/MEOW_SES-eco-5traits.Rds')
#SES_MEOW <- readRDS('FctTradeoffs/MEOW_SES-eco-5traits.Rds')
```

plots plots plots

```{r plots obs}
p_SPP <- ggplot(data = FR_MEOW_5, aes(fill = spp)) +
  geom_sf() +
  labs(fill = 'Number of Species') +
  scale_fill_viridis_c() +
  coord_sf(xlim = c(-180,180), ylim = c(-90,90)) +
  theme_minimal()

ggsave(filename = 'MEOW_SPP_map-5traits.png', plot = p_SPP, path = '~/R/thesis/FctTradeoffs', 
       width = 15, height = 10, unit = 'cm')

p_Fred_obs <- ggplot(data = FR_MEOW_5, aes(fill = fred)) +
  geom_sf() +
  labs(fill = 'FRed') +
  scale_fill_viridis_c(direction = -1) +
  coord_sf(xlim = c(-180,180), ylim = c(-90,90)) +
  theme_minimal()

ggsave(filename = 'MEOW_FRed_map-5traits.png', plot = p_Fred_obs, path = '~/R/thesis/FctTradeoffs', 
       width = 15, height = 10, unit = 'cm')

p_Fdisp_obs <- ggplot(data = FR_MEOW_5, aes(fill = fdisp)) +
  geom_sf() +
  labs(fill = 'FDisp') +
  scale_fill_viridis_c() +
  coord_sf(xlim = c(-180,180), ylim = c(-90,90)) +
  theme_minimal()

ggsave(filename = 'MEOW_FDisp_map-5traits.png', plot = p_Fdisp_obs, path = '~/R/thesis/FctTradeoffs', 
       width = 15, height = 10, unit = 'cm')

```

```{r SES plots}

p_Fred_ses <- ggplot(data = SES_MEOW, aes(fill = fred_ses)) +
  geom_sf() +
  labs(fill = 'FRed SES') +
  scale_fill_viridis_c(direction = -1) +
  coord_sf(xlim = c(-180,180), ylim = c(-90,90)) +
  theme_minimal()

ggsave(filename = 'MEOW_SES_FRed_map-eco-5traits.png', plot = p_Fred_ses, path = '~/R/thesis/FctTradeoffs', 
       width = 15, height = 10, unit = 'cm')

p_Fdisp_ses <- ggplot(data = SES_MEOW, aes(fill = fdisp_ses)) +
  geom_sf() +
  labs(fill = 'FDisp SES') +
  scale_fill_viridis_c() +
  coord_sf(xlim = c(-180,180), ylim = c(-90,90)) +
  theme_minimal()

ggsave(filename = 'MEOW_SES_FDisp_map-eco-5traits.png', plot = p_Fdisp_ses, path = '~/R/thesis/FctTradeoffs', 
       width = 15, height = 10, unit = 'cm')

#utc
mean(Fct_df$spp/Fct_df$utc) #1.046383
```

#4. Permutation test

```{r}
npermute <- 999

#need to combine Fobs and Fnull into a dataframe 

eco_obs <- data.frame(eco = rep(Fct_df$eco, each = npermute), fred_obs = rep(Fct_df$fred_obs, each = npermute), fdisp_obs = rep(Fct_df$fdisp_obs, each = npermute)) 
eco_null <- null_data  %>% 
  dplyr::select(eco, fred, fdisp) %>% 
  dplyr::rename(fred_null = fred,
                fdisp_null = fdisp) %>%
  arrange(eco)

eco_comb <- bind_cols(eco_obs, eco_null[,-"eco"])

# one-tailed permutation test
#both fred and fdisp
ptest <- eco_comb %>% 
  dplyr::mutate(fred_p_great = ifelse(fred_null > fred_obs, 1, 0),
                fred_p_less = ifelse(fred_null < fred_obs, 1, 0),
                fdisp_p_great = ifelse(fdisp_null > fdisp_obs, 1, 0),
                fdisp_p_less = ifelse(fdisp_null < fdisp_obs, 1, 0)) %>% 
  dplyr::group_by(eco) %>% 
  dplyr::summarise(fred_p_great_sum = sum(fred_p_great),
                   fred_p_less_sum = sum(fred_p_less),
                   fdisp_p_great_sum = sum(fdisp_p_great),
                   fdisp_p_less_sum = sum(fdisp_p_less)) %>% 
  dplyr::mutate(fred_p_great_test = (fred_p_great_sum + 1) / (npermute + 1),
                fred_p_less_test = (fred_p_less_sum + 1) / (npermute + 1),
                fdisp_p_great_test = (fdisp_p_great_sum + 1) / (npermute + 1),
                fdisp_p_less_test = (fdisp_p_less_sum + 1) / (npermute + 1))

ptest_results <- ptest %>% 
  dplyr::select(eco, fred_p_great_test, fred_p_less_test, fdisp_p_great_test, fdisp_p_less_test)

```


#5. Global Tradeoffs

```{r fct tradeoff}
#global tradeoff
#functional dispersion SES - functional redundancy SES
Fct_df2 <- Fct_df %>% 
  mutate(tradeoff = fdisp_ses - fred_ses) 

Fct_results <- merge(Fct_df2, ptest_results, by = 'eco')
saveRDS(Fct_results, '~/R/thesis/FctTradeoffs/Fct_analysis_results.Rds')

spp_utc <- Fct_results %>% 
  dplyr::mutate(spp_higher = ifelse(spp > utc, 1, 0),
         spp_equal = ifelse(spp == utc, 1, 0),
         spp_diff = spp-utc)

fct_red <- Fct_results %>% 
  dplyr::mutate(f_red = ifelse(fred_obs == 1, 1, 0),
                p_less = ifelse(fred_p_less_test >= 0.95, 1, 0),
                p_higher = ifelse(fred_p_great_test >= 0.95, 1, 0),
                p_disp_less = ifelse(fdisp_p_less_test >= 0.95, 1, 0),
                p_disp_higher = ifelse(fdisp_p_great_test >= 0.95, 1, 0))

fct_tr <- Fct_results %>% 
  dplyr::filter(fred_ses > 0 & fdisp_ses > 0)

spp_eq <- sum(spp_utc$spp_equal)
(spp_eq/230)*100 #24.78%
spp_high <- sum(spp_utc$spp_higher)
(spp_high/230)*100 #75.22%

fctred <- sum(fct_red$f_red)
(fctred/230)*100 #24.78%

fred_p <- sum(fct_red$p_less)
(fred_p/230)*100 #0
fred_p_high <- sum(fct_red$p_higher)
(fred_p_high/230)*100 #0
fdisp_p_l <- sum(fct_red$p_disp_less)
(fdisp_p_l/230)*100 #3.04%
fdisp_p_g <- sum(fct_red$p_disp_higher)
(fdisp_p_g/230)*100 #46.96%
```


```{r SES plots}
points_tradeoff <- ggplot(Fct_df2, aes(x = fred_ses, y = fdisp_ses, col = tradeoff)) +
  geom_point() +
  scale_color_viridis_c() +
  geom_vline(xintercept = 0, col = 'black', linetype = 'dashed') +
  geom_hline(yintercept = 0, col = 'black', linetype = 'dashed') +
  labs(x = 'Functional Redundancy SES',
       y = 'Functional Dispersion SES',
       col = 'Trade-off') +
  theme_minimal()
points_tradeoff
ggsave(filename = 'tradeoff_points-5traits.png', plot = points_tradeoff, path = '~/R/thesis/FctTradeoffs', 
       width = 7, height = 7, unit = 'cm')
tradeoff <- Fct_df2 %>% 
  dplyr::select(c('eco', 'tradeoff'))

#rename eco column to match MEOW df
colnames(tradeoff)[1] <- 'ECOREGION'
#upload MEOW df
MEOW_with_SR <- readRDS('FctTradeoffs/MEOW_with_SR.Rds')

tradeoff_MEOW <- merge(MEOW_with_SR, tradeoff, by = "ECOREGION", all.x = TRUE)

saveRDS(tradeoff_MEOW, 'FctTradeoffs/MEOW_tradeoff-eco-5traits.Rds')

p_tradeoff <- ggplot(data = tradeoff_MEOW, aes(fill = tradeoff)) +
  geom_sf() +
  labs(fill = 'Trade-off') +
  scale_fill_viridis_c() +
  coord_sf(xlim = c(-180,180), ylim = c(-90,90)) +
  theme_minimal()

p_tradeoff
ggsave(filename = 'MEOW_tradeoff_map-eco-5traits.png', plot = p_tradeoff, path = '~/R/thesis/FctTradeoffs', 
       width = 15, height = 10, unit = 'cm')

```


#6. overlay maps with MPAs
Marine protected areas (MPAs) shapefiles obtained from protectedplanet.net and stitched together into a single shp file using QGIS

```{r}
MPA_poly <- shapefile('MPA/MPAs-polygons.shp')
MPA <- st_as_sf(MPA_poly)
MPA <- MPA %>% 
  dplyr::select(WDPAID, NAME, STATUS, STATUS_YR, PARENT_ISO, NO_TK_AREA, geometry)

MPAgeo <- MPA %>% 
  dplyr::select(geometry)

saveRDS(MPAgeo, 'MPA/MPA-sf-geo.Rds')
#MPAgeo <- readRDS('MPA/MPA-sf-geo.Rds')

#FRed and MPAs
MPA_p1 <- ggplot() +
  geom_sf(data = FR_MEOW_5, aes(fill = fred)) +
  geom_sf(data = MPAgeo, col = 'red', fill = 'red') +
  labs(fill = 'FRed') +
  scale_fill_viridis_c(direction = -1) +
  coord_sf(xlim = c(-180,180), ylim = c(-90,90)) +
  theme_minimal()

ggsave(plot = MPA_p1, 'FRed+MPAs.png', width=15, height=10, units = 'cm')

#FDisp and MPA
MPA_p2 <- ggplot() +
  geom_sf(data = FR_MEOW_5, aes(fill = fdisp)) +
  geom_sf(data = MPAgeo, col = 'red', fill = 'red') +
  labs(fill = 'FDisp') +
  scale_fill_viridis_c() +
  coord_sf(xlim = c(-180,180), ylim = c(-90,90)) +
  theme_minimal()

ggsave(plot = MPA_p2, 'FDisp+MPAs.png', width=15, height=10, units = 'cm')

#FRed SES and MPA
MPA_p3 <- ggplot() +
  geom_sf(data = SES_MEOW, aes(fill = fred_ses)) +
  geom_sf(data = MPAgeo, col = 'red', fill = 'red') +
  labs(fill = 'FRed SES') +
  scale_fill_viridis_c(direction = -1) +
  coord_sf(xlim = c(-180,180), ylim = c(-90,90)) +
  theme_minimal()

ggsave(plot = MPA_p3, 'FRed_SES+MPAs.png', width=15, height=10, units = 'cm')

#FDisp SES and MPA
MPA_p4 <- ggplot() +
  geom_sf(data = SES_MEOW, aes(fill = fdisp_ses)) +
  geom_sf(data = MPAgeo, col = 'red', fill = 'red') +
  labs(fill = 'FDisp SES') +
  scale_fill_viridis_c() +
  coord_sf(xlim = c(-180,180), ylim = c(-90,90)) +
  theme_minimal()

ggsave(plot = MPA_p4, 'FDisp_SES+MPAs.png', width=15, height=10, units = 'cm')

#Tradeoff and MPA
MPA_p5 <- ggplot() +
  geom_sf(data = tradeoff_MEOW, aes(fill = tradeoff)) +
  geom_sf(data = MPAgeo, col = 'red', fill = 'red') +
  labs(fill = 'Trade-off') +
  scale_fill_viridis_c() +
  coord_sf(xlim = c(-180,180), ylim = c(-90,90)) +
  theme_minimal()

ggsave(plot = MPA_p5, 'Tradeoff+MPAs.png', width=15, height=10, units = 'cm')

```



