---
title: "Traits data"
author: "Jannika"
date: "8/8/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cetacean trait data
Gather all potentially relevant trait data from different databases:
- EltonTraits1.0: Wilman et al. 2016 https://doi.org/10.6084/m9.figshare.c.3306933.v1
- Amniote: Myhrvold et al. 2016 https://doi.org/10.6084/m9.figshare.c.3308127.v1
- Marine Megafauna Traits data from Pimiento et al. 2020 https://doi.org/10.1126/sciadv.aay7650

```{r library, message=FALSE, warning=FALSE}
library(dplyr)
library(car)
library(PVR)
library(mice)
```

#1. Upload all datasets
Datasets were subset to only include cetaceans in excel prior to uploading to avoid big files.
Furthermore, the sperm whale entry was edited manually to 'Physeter macrocephalus' in 2 of the datasets in excel to remove the synonym 'Physeter catodon' prior to uploading.

```{r data upload}
#upload first dataset
amniote <- read.csv('Traits/Amniote_Cetaceans.csv')
#list all columns and select relevant ones
colnames(amniote)
amniote <- amniote %>% 
  dplyr::select(c(4:5, 8:16, 20:24, 34)) %>% 
  #genus and species name are in 2 columns - merge them
  tidyr::unite(species, c(genus, species)) 
#seperate the species from genus names with a space to mimic other datasets
amniote$species <- sub('_',' ',amniote$species)

#upload second dataset
elton <- read.csv('Traits/EltonTraits_cetaceans.csv')
#list all columns and select relevant ones
colnames(elton)
elton <- elton %>% 
  dplyr::select(c(2, 4:13, 19:21, 24)) %>% 
  dplyr::rename(species = Scientific)

#upload third dataset
pimiento <- read.csv('Traits/Pimiento-megafauna_whales.csv')
#list all columns and select relevant ones
colnames(pimiento)
pimiento <- pimiento %>% 
  dplyr::select(c(1,6:8, 10:13, 15:24)) 
  
#merge datasets
traits1 <- merge(amniote, elton, by = 'species', all.x = TRUE, all.y = TRUE)
traits <- merge(traits1, pimiento, by = 'species', all.x = TRUE, all.y = TRUE)
```

#2. Clean dataset
1. remove traits with species coverage of less than 50%
2. calculate mean body mass across all 3 estimates from the different datasets and drop the original columns
3. remove diet categories from EltonTraits1.0 that aren't applicable
4. remove any species which aren't marine
5. calculate habitat breadth across the 7 ocean basins and remove the 7 columns listing the presence in each ocean basin
6. convert diel activity into a single column with 5 numerical categories (nocturnal = 1, nocturnal + twilight = 2, nocturnal + twilight + diurnal = 3, twilight + diurnal = 4, diurnal = 5) and drop the original 3 columns
7. convert categorical varibales from Piminetos dataset into numerical categories (migration: yes = 1, no = 0; breeding_site: coastal = 0, oceanic = 1, both = 2; vertical: pelagic = 0, benthopelagic = 1, benthic = 2; diet: plankton = 0, inverts = 1, inverts/fish = 2, fish = 3, fish/higherverts = 4; feeding: filter = 0, predator = 1; group_behaviour: solitary = 0, social = 1, both = 2)

```{r clean datasets}
#1. replace empty entries and -999 entries with NA
traits[traits == ''] <- NA
traits[traits == '-999'] <- NA

#2. check how many NAs are in each column
#total 90 observations - remove columns with more than 50% NAs (>= 45)
NAs <- colSums(is.na(traits))
NAs <- as.data.frame(NAs)
#remove columns with 50% NAs
traits2 <- traits[, which(colMeans(!is.na(traits)) > 0.5)] #remove 4 columns

#average bodymass from all 3 datasets, ignore NAs and remove original weight columns
traits2 <- traits2 %>% 
  dplyr::mutate(meanBodyMass = rowMeans(.[c('weight', 'BodyMass.Value', 'adult_body_mass_g')], na.rm = T)) %>% 
  dplyr::select(-c('weight', 'BodyMass.Value', 'adult_body_mass_g')) %>% 
#remove diet categories which do not make sense 
  dplyr::select(-c('Diet.Vect', 'Diet.Vunk', 'Diet.Scav', 'Diet.Fruit', 'Diet.Nect', 'Diet.Seed', 'Diet.PlantO'))

#remove freshwater species
#upload marine_species_only data
marine <- read.csv('Traits/marine_only_species.csv')
colnames(marine)
marine <- dplyr::rename(marine, species = ï..species)
colnames(marine)
#merge marine species with traits data and drop all rows that are not included in the marine species list
traits3 <- merge(marine, traits2, by = 'species')

#make habitat breadth category and drop the original ones
traits4 <- traits3 %>% 
  dplyr::mutate(HabitatBreadth = rowSums(.[c('Arctic', 'North.Pacific', 'South.Pacific', 'North.Atlantic', 'South.Atlantic', 'Indian', 'Southern')], na.rm = T)) %>% 
  dplyr::select(-c('Arctic', 'North.Pacific', 'South.Pacific', 'North.Atlantic', 'South.Atlantic', 'Indian', 'Southern')) %>% 

  #make diel activity category and drop the original ones
  #nocturnal = 1, noct+twilight = 2, all = 3, diurnal + twilight = 4, diurnal = 5
  dplyr::mutate(DielActivity = 
            if_else(Activity.Nocturnal == 1 & Activity.Crepuscular == 0 & Activity.Diurnal == 0, 1,
            if_else(Activity.Nocturnal == 1 & Activity.Crepuscular == 1 & Activity.Diurnal == 0, 2, 
            if_else(Activity.Nocturnal == 1 & Activity.Crepuscular == 1 & Activity.Diurnal == 1, 3,
            if_else(Activity.Nocturnal == 0 & Activity.Crepuscular == 1 & Activity.Diurnal == 1, 4, 5))))) %>% 
  dplyr::select(-c('Activity.Nocturnal', 'Activity.Crepuscular', 'Activity.Diurnal'))

#make all categorical variables into numerical ones
traits5 <- traits4 %>% 
  dplyr::mutate(migration = if_else(migration == 'marine', 1, 0)) %>% 
  dplyr::mutate(breeding_site = if_else(breeding_site == 'coastal', 0, 
                                        if_else(breeding_site == 'oceanic', 1, 2))) %>% 
  dplyr::mutate(vertical = if_else(vertical == 'pelagic', 0,
                                   if_else(vertical == 'benthopelagic', 1, 2))) %>% 
  dplyr::mutate(diet = if_else(diet == 'plankton', 0, 
                               if_else(diet == 'inverts', 1,
                                       if_else(diet == 'inverts/fish', 2,
                                               if_else(diet == 'fish', 3, 4))))) %>% 
  dplyr::mutate(feeding = if_else(feeding == 'filter-feeding', 0, 1)) %>% 
  dplyr::mutate(group_behaviour = if_else(group_behaviour == 'solitary', 0,
                                          if_else(group_behaviour == 'social', 1, 2)))


```

#3. Regression analysis
Reduce multicollinearity by regressing each trait against all other traits using linear regression models.

```{r trait selection}
#reduce multicollinearity by regressing each trait against all other traits using regression
#doesn't work if there are many NAs in the dataset thus subset for correlation by removing NAs
traitsNA <- traits4 %>% 
  na.omit()
#do linear regression models of all numerical variables first (not ones I made numerical)- check if VIF is under 5 then later do under 3
mod1 <- lm(female_maturity_d ~ litter_or_clutch_size_n + litters_or_clutches_per_y +
           maximum_longevity_y + gestation_d + weaning_d + birth_or_hatching_weight_g +
           longevity_y + male_maturity_d + inter_litter_or_interbirth_interval_y +
            meanBodyMass + HabitatBreadth + DielActivity, data = traitsNA) 
vif(mod1) #only litter size and diel activity under 3, habitat breadth under 5 all others over 5

mod2 <- lm(litter_or_clutch_size_n ~ litters_or_clutches_per_y +
           maximum_longevity_y + gestation_d + weaning_d + birth_or_hatching_weight_g +
           longevity_y + male_maturity_d + inter_litter_or_interbirth_interval_y +
            meanBodyMass + HabitatBreadth + DielActivity + female_maturity_d, data = traitsNA) 
vif(mod2) #only diel activity under 3, habitat under 5

mod3 <- lm(litters_or_clutches_per_y  ~ maximum_longevity_y + gestation_d + weaning_d +
             birth_or_hatching_weight_g + longevity_y + male_maturity_d +
             inter_litter_or_interbirth_interval_y + meanBodyMass + HabitatBreadth + DielActivity +
             female_maturity_d + litter_or_clutch_size_n, data = traitsNA) 
vif(mod3) #only litter size and diel activity under 3, weaning and habitat breadth under 5

mod4 <- lm(maximum_longevity_y ~ gestation_d + weaning_d +
             birth_or_hatching_weight_g + longevity_y + male_maturity_d +
             inter_litter_or_interbirth_interval_y + meanBodyMass + HabitatBreadth + DielActivity +
             female_maturity_d + litter_or_clutch_size_n + litters_or_clutches_per_y, data = traitsNA) 
vif(mod4) #litter size and diel activity under 3, habitat breadth under 5

mod5 <- lm(gestation_d ~ weaning_d +
             birth_or_hatching_weight_g + longevity_y + male_maturity_d +
             inter_litter_or_interbirth_interval_y + meanBodyMass + HabitatBreadth + DielActivity +
             female_maturity_d + litter_or_clutch_size_n + litters_or_clutches_per_y +
              maximum_longevity_y, data = traitsNA) 
vif(mod5) #litter size and diel activity and weaning under 3, habitat breadth under 5

mod6 <- lm(weaning_d ~ birth_or_hatching_weight_g + longevity_y + male_maturity_d +
             inter_litter_or_interbirth_interval_y + meanBodyMass + HabitatBreadth + DielActivity +
             female_maturity_d + litter_or_clutch_size_n + litters_or_clutches_per_y +
              maximum_longevity_y + gestation_d, data = traitsNA) 
vif(mod6) #litter size and diel activity under 3, habitat breadth and gestation under 5

mod7 <- lm(birth_or_hatching_weight_g ~ longevity_y + male_maturity_d +
             inter_litter_or_interbirth_interval_y + meanBodyMass + HabitatBreadth + DielActivity +
             female_maturity_d + litter_or_clutch_size_n + litters_or_clutches_per_y +
              maximum_longevity_y + gestation_d + weaning_d, data = traitsNA) 
vif(mod7) #litter size and diel activity under 3, bodymass and habitat breadth under 5

mod8 <- lm(longevity_y ~ male_maturity_d +
             inter_litter_or_interbirth_interval_y + meanBodyMass + HabitatBreadth + DielActivity +
             female_maturity_d + litter_or_clutch_size_n + litters_or_clutches_per_y +
              maximum_longevity_y + gestation_d + weaning_d +
              birth_or_hatching_weight_g, data = traitsNA) 
vif(mod8) #habitat breadth, diel activity, litter size under 3, gestation and weaning under 5

mod9 <- lm(male_maturity_d ~
             inter_litter_or_interbirth_interval_y + meanBodyMass + HabitatBreadth + DielActivity +
             female_maturity_d + litter_or_clutch_size_n + litters_or_clutches_per_y +
              maximum_longevity_y + gestation_d + weaning_d +
              birth_or_hatching_weight_g + longevity_y, data = traitsNA) 
vif(mod9) #habitat breadth, diel activity, litter size under 3, weaning under 5

mod10 <- lm(inter_litter_or_interbirth_interval_y ~ meanBodyMass + HabitatBreadth + DielActivity +
             female_maturity_d + litter_or_clutch_size_n + litters_or_clutches_per_y +
              maximum_longevity_y + gestation_d + weaning_d +
              birth_or_hatching_weight_g + longevity_y + male_maturity_d, data = traitsNA) 
vif(mod10) #diel activity, litter size under 3, habitat breadth under 5

mod11 <- lm(meanBodyMass ~ HabitatBreadth + DielActivity +
             female_maturity_d + litter_or_clutch_size_n + litters_or_clutches_per_y +
              maximum_longevity_y + gestation_d + weaning_d +
              birth_or_hatching_weight_g + longevity_y + male_maturity_d +
               inter_litter_or_interbirth_interval_y, data = traitsNA) 
vif(mod11) #diel activity, litter size under 3, habitat breadth under 5

mod12 <- lm(HabitatBreadth ~ DielActivity +
             female_maturity_d + litter_or_clutch_size_n + litters_or_clutches_per_y +
              maximum_longevity_y + gestation_d + weaning_d +
              birth_or_hatching_weight_g + longevity_y + male_maturity_d +
               inter_litter_or_interbirth_interval_y + meanBodyMass, data = traitsNA) 
vif(mod12) #diel activity, litter size under 3, weaning under 5

mod13 <- lm(DielActivity ~ female_maturity_d + litter_or_clutch_size_n + 
              litters_or_clutches_per_y + maximum_longevity_y + gestation_d + weaning_d +
              birth_or_hatching_weight_g + longevity_y + male_maturity_d +
              inter_litter_or_interbirth_interval_y + meanBodyMass + HabitatBreadth, data = traitsNA) 
vif(mod13) #litter_or_clutch size under 3, habitat breadth, weaning under 5

#select all of the traits that were under 5 in any of the models
traits_vif5 <- traitsNA %>% 
  select(c('litter_or_clutch_size_n', 'DielActivity', 'HabitatBreadth', 'weaning_d', 'gestation_d', 'meanBodyMass'))
#check again by regression
mod_1 <- lm(litter_or_clutch_size_n ~ DielActivity + HabitatBreadth + weaning_d +
              gestation_d + meanBodyMass, data = traits_vif5)
vif(mod_1) #all under 3

#now add the remaining traits to the regression models
#I converted the categorical ones into numerical categories so they can also be the dependent varibable
traitsNA2 <- traits5 %>% 
  na.omit()
#don't use feeding as it is correlated to diet
cor(traitsNA2$diet, traitsNA2$feeding) #0.79
mod.1 <- lm(litter_or_clutch_size_n ~ DielActivity + HabitatBreadth + weaning_d +
              gestation_d + meanBodyMass + breeding_site + migration + vertical + 
              diet + group_behaviour, data = traitsNA2)
vif(mod.1) #weaining and group_behaviour OVER 3

mod.2 <- lm(DielActivity ~ HabitatBreadth + weaning_d +
              gestation_d + meanBodyMass + breeding_site + migration + vertical + 
              diet + group_behaviour + litter_or_clutch_size_n, data = traitsNA2)
vif(mod.2) #weaning and group_behaviour OVER 3

mod.3 <- lm(HabitatBreadth ~ weaning_d + gestation_d + meanBodyMass + breeding_site + 
              migration + vertical + diet + group_behaviour + 
              litter_or_clutch_size_n + DielActivity, data = traitsNA2)
vif(mod.3) #weaning and group_behaviour OVER 3

mod.4 <- lm(weaning_d ~ gestation_d + meanBodyMass + breeding_site + 
              migration + vertical + diet + group_behaviour + 
              litter_or_clutch_size_n + DielActivity + HabitatBreadth, data = traitsNA2)
vif(mod.4) #group_behaviour OVER 3 

mod.5 <- lm(gestation_d ~ meanBodyMass + breeding_site + migration + vertical + 
              diet + group_behaviour + litter_or_clutch_size_n + DielActivity + 
              HabitatBreadth + weaning_d, data = traitsNA2)
vif(mod.5) #group_behaviour and weaning OVER 3

mod.6 <- lm(meanBodyMass ~ breeding_site + migration + vertical + 
              diet + group_behaviour + litter_or_clutch_size_n + DielActivity + 
              HabitatBreadth + weaning_d + gestation_d, data = traitsNA2)
vif(mod.6) #group_behaviour and weaning OVER 3

mod.7 <- lm(breeding_site ~ migration + vertical + diet + group_behaviour + 
              litter_or_clutch_size_n + DielActivity + HabitatBreadth + 
              weaning_d + gestation_d + meanBodyMass, data = traitsNA2)
vif(mod.7) #group_behaviour and weaning OVER 3

mod.8 <- lm(migration ~ vertical + diet + group_behaviour + 
              litter_or_clutch_size_n + DielActivity + HabitatBreadth + 
              weaning_d + gestation_d + meanBodyMass + breeding_site, data = traitsNA2)
vif(mod.8) #group_behaviour and weaning OVER 3

mod.9 <- lm(vertical ~ diet + group_behaviour + 
              litter_or_clutch_size_n + DielActivity + HabitatBreadth + 
              weaning_d + gestation_d + meanBodyMass + breeding_site + migration, data = traitsNA2)
vif(mod.9) #group_behaviour and weaning OVER 3

mod.10 <- lm(diet ~ group_behaviour + litter_or_clutch_size_n + DielActivity + 
               HabitatBreadth + weaning_d + gestation_d + meanBodyMass + 
               breeding_site + migration + vertical, data = traitsNA2)
vif(mod.10) #group_behaviour and weaning OVER 3

mod.11 <- lm(group_behaviour ~ litter_or_clutch_size_n + DielActivity + 
               HabitatBreadth + weaning_d + gestation_d + meanBodyMass + 
               breeding_site + migration + vertical + diet, data = traitsNA2)
vif(mod.11) #weaning OVER 3

#remove weaning and group behaviour!
traits_vif3 <- traitsNA2 %>% 
  select(c('litter_or_clutch_size_n', 'DielActivity', 'HabitatBreadth', 'gestation_d',
           'meanBodyMass', 'breeding_site', 'migration', 'vertical', 'diet'))

mod_2 <- lm(litter_or_clutch_size_n ~ DielActivity + HabitatBreadth + gestation_d + 
                meanBodyMass + breeding_site + migration + vertical + diet, data = traits_vif3)
vif(mod_2)
#all under 3
```

#subset trait data
Subset trait dataset according to the results from the regression analysis and save these.

```{r trait subset}
#generate final list with traits to use for further analysis and save the files as .csv
traits_final <- traits5 %>% 
  select(c('species', 'litter_or_clutch_size_n', 'gestation_d', 'DielActivity', 'HabitatBreadth',
           'meanBodyMass', 'breeding_site', 'migration', 'vertical', 'diet')) %>% 
  rename(LitterSize_n = litter_or_clutch_size_n)
write.csv(traits_final, 'Traits/traits.csv')
#traits_final <- read.csv('Traits/traits.csv')

#make a file for family and species
family <- traits5 %>% 
  select(c('species', 'family'))

write.csv(family, 'Traits/family.csv')
#make an iucn file
iucn <- traits5 %>% 
  select(c('species', 'IUCN'))

write.csv(iucn, 'Traits/iucn.csv')

```

##Trait imputation
Following method from Cooke et al. 2019 https://doi.org/10.1038/s41467-019-10284-z

#1. Upload trait data

```{r trait setup}
#set up trait data
trait <- read.csv('Traits/traits.csv')
trait <- trait %>% 
  dplyr::select(-X)

family <- read.csv('Traits/family.csv')
family <- family %>% 
  dplyr::select(-X)

tr <- trait %>% 
  # add class data
  dplyr::left_join(family, by = "species")
```

#2. Upload phylogenetic data
phylogentic tree cetacean subset from MammalTree obtained from http://vertlife.org/phylosubsets/ (Upham et al 2019 'Inferring the mammal tree: Species-level sets of phylogenies for questions in ecology, evolution, and conservation.')

```{r phylogenetic tree}
# phylogenetic data
ceta_tree <- read.nexus('~/Sussex/thesis/Traits data/CetaceaTree/output.nex')
class(ceta_tree)
#computes the maximum clade credibility tree from a sample of trees
tree <- phangorn::maxCladeCred(ceta_tree)
PVR <- PVRdecomp(tree)
#select out species names in the order of the phylo trees
species <- PVR@phylo$tip.label
species <- as.data.frame(species)
#change species names
species$species <- sub("_", " ", species$species)
#make a dataframe with the phylogenetic eigenvectors and select the first 10 
eigen <- as.data.frame(PVR@Eigen$vectors)
eigen10 <- eigen %>% 
  select('c1':'c10')
#combine phylogenetic eigenvectors to species names and rename column names
pem <- cbind(species, eigen10)
oldnames <- c('c1', 'c2', 'c3', 'c4', 'c5', 'c6', 'c7', 'c8', 'c9', 'c10')
newnames <- c('V_1', 'V_2', 'V_3', 'V_4', 'V_5', 'V_6', 'V_7', 'V_8', 'V_9', 'V_10')
pem <- pem %>% 
  rename_at(vars(oldnames), ~ newnames)

#save phylogenetic eigenvectors dataframe
saveRDS(pem, 'Traits/pem_cetaceans.Rds')
pem <- readRDS('Traits/pem_cetaceans.Rds')
```


```{r imputation}
#sort both data tables by binomial alphabetically
tr_sort <- tr[order(tr[,'species']), ]
pem_sort <- pem[order(pem[, 'species']),]
# join trait and phylogenetic data
tr_ceta <- dplyr::left_join(tr_sort, pem_sort, by = c('species'))

#change dataframe to get rid off errors - try making the variables more similar...
tr_ceta1 <- tr_ceta %>% 
  dplyr::mutate(meanBodyMass = meanBodyMass/1000000)

# missing data pattern
mice::md.pattern(dplyr::select(tr_ceta1, -species, -family, -c(V_1:V_10)))

# run multiple imputation
tr_mi_raw <- mice(tr_ceta1, m = 25, maxit = 100, seed = 20)
# method = pmm predictive mean matching

# summary of multiple imputation results including predictor matrix (which variables were used to predict missing values)
summary(tr_mi_raw)

# save: tr_mi_raw
saveRDS(tr_mi_raw, 'Traits/tr_mi_raw.rds')
# dataframe of raw multiple imputated trait data for mammals (data.frame)

tr_mi <- lapply(1:25, function(x) {
  out <- mice::complete(tr_mi_raw, action = x) %>% 
    dplyr::select(species:diet, -family) 
})

# save: tr_mi
saveRDS(tr_mi, 'Traits/tr_mi.rds')

```


