---
title: "Traits data"
author: "Jannika"
date: "8/8/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cetacean trait data
Gather all potentially relevant trait data from different databases:
- EltonTraits1.0: Wilman et al. 2016 https://doi.org/10.6084/m9.figshare.c.3306933.v1
- Amniote: Myhrvold et al. 2016 https://doi.org/10.6084/m9.figshare.c.3308127.v1
- Marine Megafauna Traits data from Pimiento et al. 2020 https://doi.org/10.1126/sciadv.aay7650

```{r library, message=FALSE, warning=FALSE}
library(dplyr)
library(car)
library(DescTools)
library(PVR)
library(mice)
```

#1. Upload all datasets
Datasets were subset to only include cetaceans in excel prior to uploading to avoid big files.
Furthermore, the sperm whale entry was edited manually to 'Physeter macrocephalus' in 2 of the datasets in excel to remove the synonym 'Physeter catodon' and the spelling of 'Mesoplodon ginkgodens' was corrected in the Marine Megafauna Traits dataset prior to uploading.

```{r data upload}
#upload first dataset
amniote <- read.csv('Traits/Amniote_Cetaceans.csv')
#list all columns and select relevant ones
colnames(amniote)
amniote <- amniote %>% 
  dplyr::select(c(4:5, 8:16, 20:24, 34)) %>% 
  #genus and species name are in 2 columns - merge them
  tidyr::unite(species, c(genus, species)) 
#seperate the species from genus names with a space to mimic other datasets
amniote$species <- sub('_',' ',amniote$species)

#upload second dataset
elton <- read.csv('Traits/EltonTraits_cetaceans.csv')
#list all columns and select relevant ones
colnames(elton)
elton <- elton %>% 
  dplyr::select(c(2, 4:13, 19:21, 24)) %>% 
  dplyr::rename(species = Scientific)

#upload third dataset
pimiento <- read.csv('Traits/Pimiento-megafauna_whales.csv')
#list all columns and select relevant ones
colnames(pimiento)
pimiento <- pimiento %>% 
  dplyr::select(c(1,6:8, 10:13, 15:24)) 
  
#merge datasets
traits1 <- merge(amniote, elton, by = 'species', all.x = TRUE, all.y = TRUE)
traits <- merge(traits1, pimiento, by = 'species', all.x = TRUE, all.y = TRUE)
```

#2. Clean dataset
1. remove traits with species coverage of less than 50%
2. calculate mean body mass across all 3 estimates from the different datasets and drop the original columns
3. remove diet categories from EltonTraits1.0 that aren't applicable
4. remove any species which aren't marine
5. calculate habitat breadth across the 7 ocean basins and remove the 7 columns listing the presence in each ocean basin
6. convert diel activity into a single column with 5 numerical categories (nocturnal = 1, nocturnal + twilight = 2, nocturnal + twilight + diurnal = 3, twilight + diurnal = 4, diurnal = 5) and drop the original 3 columns
7. convert categorical varibales from Piminetos dataset into numerical categories (migration: yes = 1, no = 0; breeding_site: coastal = 0, oceanic = 1, both = 2; vertical: pelagic = 0, benthopelagic = 1, benthic = 2; diet: plankton = 0, inverts = 1, inverts/fish = 2, fish = 3, fish/higherverts = 4; feeding: filter = 0, predator = 1; group_behaviour: solitary = 0, social = 1, both = 2)

```{r clean datasets}
#1. replace empty entries and -999 entries with NA
traits[traits == ''] <- NA
traits[traits == '-999'] <- NA

#2. check how many NAs are in each column
#total 90 observations - remove columns with more than 50% NAs (>= 45)
NAs <- colSums(is.na(traits))
NAs <- as.data.frame(NAs)
#remove columns with 50% NAs
traits2 <- traits[, which(colMeans(!is.na(traits)) > 0.5)] #remove 4 columns

#average bodymass from all 3 datasets, ignore NAs and remove original weight columns
traits2 <- traits2 %>% 
  dplyr::mutate(meanBodyMass = rowMeans(.[c('weight', 'BodyMass.Value', 'adult_body_mass_g')], na.rm = T)) %>% 
  dplyr::select(-c('weight', 'BodyMass.Value', 'adult_body_mass_g')) %>% 
#remove diet categories which do not make sense 
  dplyr::select(-c('Diet.Vect', 'Diet.Vunk', 'Diet.Scav', 'Diet.Fruit', 'Diet.Nect', 
                   'Diet.Seed', 'Diet.PlantO')) %>% 
  #remove litter and clutch size n as it does not make sense when looking at this single taxa with very similar values
  dplyr::select(-c('litter_or_clutch_size_n'))

#remove freshwater species
#upload marine_species_only data
marine <- read.csv('Traits/marine_only_species.csv')
colnames(marine)
marine <- dplyr::rename(marine, species = ï..species)
colnames(marine)
#merge marine species with traits data and drop all rows that are not included in the marine species list
traits3 <- merge(marine, traits2, by = 'species')

#make habitat breadth category and drop the original ones
traits4 <- traits3 %>% 
  dplyr::mutate(HabitatBreadth = rowSums(.[c('Arctic', 'North.Pacific', 'South.Pacific', 'North.Atlantic', 'South.Atlantic', 'Indian', 'Southern')], na.rm = T)) %>% 
  dplyr::select(-c('Arctic', 'North.Pacific', 'South.Pacific', 'North.Atlantic', 'South.Atlantic', 'Indian', 'Southern')) %>% 
  #make diel activity category and drop the original ones
  dplyr::mutate(DielActivity = 
            if_else(Activity.Nocturnal == 1 & Activity.Crepuscular == 0 & Activity.Diurnal == 0, 'N',
            if_else(Activity.Nocturnal == 0 & Activity.Crepuscular == 1 & Activity.Diurnal == 0, 'C',
            if_else(Activity.Nocturnal == 0 & Activity.Crepuscular == 0 & Activity.Diurnal == 1, 'D', 
            if_else(Activity.Nocturnal == 1 & Activity.Crepuscular == 1 & Activity.Diurnal == 0, 'NC',
            if_else(Activity.Nocturnal == 0 & Activity.Crepuscular == 1 & Activity.Diurnal == 1, 'CD', 
            if_else(Activity.Nocturnal == 1 & Activity.Crepuscular == 0 & Activity.Diurnal == 1, 'ND',
            if_else(Activity.Nocturnal == 1 & Activity.Crepuscular == 1 & Activity.Diurnal == 1, 'NCD',
                    'None')))))))) %>% 
  dplyr::select(-c('Activity.Nocturnal', 'Activity.Crepuscular', 'Activity.Diurnal')) %>%
  #set all categorical fariables as factor and habitat breadth as discrete variable (integer)
  mutate(DielActivity = as.factor(DielActivity),
         habitat = as.factor(habitat),
         breeding_site = as.factor(breeding_site),
         migration = as.factor(migration),
         vertical = as.factor(vertical),
         diet = as.factor(diet),
         feeding = as.factor(feeding),
         group_behaviour = as.factor(group_behaviour),
         HabitatBreadth = as.integer(HabitatBreadth))

```

#3. Regression analysis
Reduce multicollinearity by regressing each trait against all other traits using linear regression models.

```{r trait selection continuous}
#reduce multicollinearity by regressing each continuous trait against all other continuous traits using regression
#doesn't work if there are many NAs in the dataset thus subset for correlation by removing NAs
traitsNA <- traits4 %>% 
  na.omit()
#do linear regression models and check if VIF is under 5 then later do under 3
mod1 <- lm(female_maturity_d ~ litters_or_clutches_per_y + maximum_longevity_y + 
             gestation_d + weaning_d + birth_or_hatching_weight_g + longevity_y + 
             male_maturity_d + inter_litter_or_interbirth_interval_y + 
             meanBodyMass + HabitatBreadth, data = traitsNA) 
vif(mod1) #habitat breadth under 3, weaning under 5

mod2 <- lm(litters_or_clutches_per_y  ~ maximum_longevity_y + gestation_d + weaning_d +
             birth_or_hatching_weight_g + longevity_y + male_maturity_d +
             inter_litter_or_interbirth_interval_y + meanBodyMass + HabitatBreadth + 
             female_maturity_d, data = traitsNA) 
vif(mod2) #none under 3, weaning, inter litter interval y and habitat breadth under 5

mod3 <- lm(maximum_longevity_y ~ gestation_d + weaning_d +
             birth_or_hatching_weight_g + longevity_y + male_maturity_d +
             inter_litter_or_interbirth_interval_y + meanBodyMass + HabitatBreadth + 
             female_maturity_d + litters_or_clutches_per_y, data = traitsNA) 
vif(mod3) #none under 3, habitat breadth and weaning under 5

mod4 <- lm(gestation_d ~ weaning_d +
             birth_or_hatching_weight_g + longevity_y + male_maturity_d +
             inter_litter_or_interbirth_interval_y + meanBodyMass + HabitatBreadth + 
             female_maturity_d + litters_or_clutches_per_y + maximum_longevity_y, data = traitsNA) 
vif(mod4) #habitat breadth and weaning under 3, none under 5

mod5 <- lm(weaning_d ~ birth_or_hatching_weight_g + longevity_y + male_maturity_d +
             inter_litter_or_interbirth_interval_y + meanBodyMass + HabitatBreadth +
             female_maturity_d + litters_or_clutches_per_y +
              maximum_longevity_y + gestation_d, data = traitsNA) 
vif(mod5) #habitat breadth under 3, gestation under 5

mod6 <- lm(birth_or_hatching_weight_g ~ longevity_y + male_maturity_d +
             inter_litter_or_interbirth_interval_y + meanBodyMass + HabitatBreadth +
             female_maturity_d + litters_or_clutches_per_y +
              maximum_longevity_y + gestation_d + weaning_d, data = traitsNA) 
vif(mod6) #none under 3, bodymass, habitat breadth and weaning under 5

mod7 <- lm(longevity_y ~ male_maturity_d +
             inter_litter_or_interbirth_interval_y + meanBodyMass + HabitatBreadth +
             female_maturity_d + litters_or_clutches_per_y +
              maximum_longevity_y + gestation_d + weaning_d +
              birth_or_hatching_weight_g, data = traitsNA) 
vif(mod7) #habitat breadth under 3, gestation and weaning under 5

mod8 <- lm(male_maturity_d ~
             inter_litter_or_interbirth_interval_y + meanBodyMass + HabitatBreadth +
             female_maturity_d + litters_or_clutches_per_y +
              maximum_longevity_y + gestation_d + weaning_d +
              birth_or_hatching_weight_g + longevity_y, data = traitsNA) 
vif(mod8) #habitat breadth under 3, weaning under 5

mod9 <- lm(inter_litter_or_interbirth_interval_y ~ meanBodyMass + HabitatBreadth +
             female_maturity_d + litters_or_clutches_per_y + maximum_longevity_y + 
             gestation_d + weaning_d + birth_or_hatching_weight_g + longevity_y + 
             male_maturity_d, data = traitsNA) 
vif(mod9) #none under 3, habitat breadth, litters per year and weaning under 5

mod10 <- lm(meanBodyMass ~ HabitatBreadth +
             female_maturity_d + litters_or_clutches_per_y +
              maximum_longevity_y + gestation_d + weaning_d +
              birth_or_hatching_weight_g + longevity_y + male_maturity_d +
               inter_litter_or_interbirth_interval_y, data = traitsNA) 
vif(mod10) #none under 3, habitat breadth and weaning under 5

mod11 <- lm(HabitatBreadth ~ female_maturity_d + litters_or_clutches_per_y + 
              maximum_longevity_y + gestation_d + weaning_d +
              birth_or_hatching_weight_g + longevity_y + male_maturity_d +
               inter_litter_or_interbirth_interval_y + meanBodyMass, data = traitsNA) 
vif(mod11) #none under 3, weaning under 5

#select all of the traits that were under 5 in any of the models
traits_vif5 <- traitsNA %>% 
  select(c('litters_or_clutches_per_y', 'inter_litter_or_interbirth_interval_y', 
           'HabitatBreadth', 'weaning_d', 'gestation_d', 'meanBodyMass'))
#check again by regression
mod_1 <- lm(litters_or_clutches_per_y ~ inter_litter_or_interbirth_interval_y +
              HabitatBreadth + weaning_d + gestation_d + meanBodyMass, data = traits_vif5)
vif(mod_1) #all under 3

mod_2 <- lm(inter_litter_or_interbirth_interval_y ~
              HabitatBreadth + weaning_d + gestation_d + meanBodyMass +
              litters_or_clutches_per_y, data = traits_vif5)
vif(mod_2) #all under 3

mod_3 <- lm(HabitatBreadth ~ weaning_d + gestation_d + meanBodyMass +
              litters_or_clutches_per_y + inter_litter_or_interbirth_interval_y, data = traits_vif5)
vif(mod_3) #litter or clutches per year over 3

mod_4 <- lm(weaning_d ~ gestation_d + meanBodyMass + litters_or_clutches_per_y +
              inter_litter_or_interbirth_interval_y + HabitatBreadth, data = traits_vif5)
vif(mod_4) #litter or clutches per year over 3

mod_5 <- lm(gestation_d ~ meanBodyMass +
              litters_or_clutches_per_y + inter_litter_or_interbirth_interval_y +
              HabitatBreadth + weaning_d, data = traits_vif5)
vif(mod_5) #litter or clutches per year over 3

mod_6 <- lm(meanBodyMass ~ litters_or_clutches_per_y + inter_litter_or_interbirth_interval_y +
              HabitatBreadth + weaning_d + gestation_d, data = traits_vif5)
vif(mod_6) #litter or clutches per year over 3

traits_vif3_cont <- traits_vif5 %>% 
  select(-c('litters_or_clutches_per_y'))

mod.1 <- lm(inter_litter_or_interbirth_interval_y ~
              HabitatBreadth + weaning_d + gestation_d + meanBodyMass, data = traits_vif3_cont)
vif(mod.1) #all under 3
mod.2 <- lm(HabitatBreadth ~ weaning_d + gestation_d + meanBodyMass +
              inter_litter_or_interbirth_interval_y, data = traits_vif3_cont)
vif(mod.2) #all under 3
mod.3 <- lm(weaning_d ~ gestation_d + meanBodyMass +
              inter_litter_or_interbirth_interval_y +
              HabitatBreadth, data = traits_vif3_cont)
vif(mod.3) #all under 3
mod.4 <- lm(gestation_d ~ meanBodyMass +
              inter_litter_or_interbirth_interval_y +
              HabitatBreadth + weaning_d, data = traits_vif3_cont)
vif(mod.4) #all under 3
mod.5 <- lm(meanBodyMass ~ inter_litter_or_interbirth_interval_y +
              HabitatBreadth + weaning_d + gestation_d, data = traits_vif3_cont)
vif(mod.5) #all under 3

vif3 <- cor(traits_vif3_cont)
#remove weaning as it is medium correlated with gestation (> 0.5)
traits_vif3_cont <- traits_vif3_cont %>% 
  select(-weaning_d)

```

```{r trait selection categorical traits}
#do chi-square test to test correlation between 2 categorical traits, H0 assume they are independent, if p < 0.05 reject null hypothesis that they are independent thus they are correlated 

chisq.test(traitsNA$breeding_site, traitsNA$migration)
#p-value: 0.5796 - indep
chisq.test(traitsNA$breeding_site, traitsNA$vertical)
#p-value: 0.8131 - indep
chisq.test(traitsNA$breeding_site, traitsNA$diet)
#p-value: 0.8215 - indep
chisq.test(traitsNA$breeding_site, traitsNA$feeding)
#p-value: 0.1433 - indep
chisq.test(traitsNA$breeding_site, traitsNA$group_behaviour)
#p-value: 0.1069 - indep

chisq.test(traitsNA$migration, traitsNA$vertical)
#p-value: 0.7245 - indep
chisq.test(traitsNA$migration, traitsNA$diet)
#p-value: 0.8347 - indep
chisq.test(traitsNA$migration, traitsNA$feeding)
#p-value: 1 - correlated
chisq.test(traitsNA$migration, traitsNA$group_behaviour)
#p-value: 0.2171 - indep

chisq.test(traitsNA$vertical, traitsNA$diet)
#p-value: 0.2802 - indep
chisq.test(traitsNA$vertical, traitsNA$feeding)
#p-value: 0.0344 - dependent
chisq.test(traitsNA$vertical, traitsNA$group_behaviour)
#p-value: 0.6052 - indep

chisq.test(traitsNA$diet, traitsNA$feeding)
#p-value: 7.987e-05 - dependent
chisq.test(traitsNA$diet, traitsNA$group_behaviour)
#p-value: 0.01263 - dependent

chisq.test(traitsNA$feeding, traitsNA$group_behaviour)
#p-value: 0.001028 - dependent


#calculate Cramer V and Pearsons Correlation Coefficient C for all 6 categorical traits, pairwise with symmetry = true and C corrected (correct for sample size, standardize results to lie between 0 and 1)
V <- PairApply(traitsNA[,c('DielActivity', 'breeding_site','migration','vertical', 'diet', 'feeding', 'group_behaviour')], CramerV, symmetric = TRUE)

#PairApply(traitsNA[,c("breeding_site","migration","vertical", 'diet', 'feeding', 'group_behaviour')], ContCoef, symmetric = TRUE, correct = TRUE)

#select categorical traits with Cramer V coefficient below 0.5 as a medium correlation cutoff 
traitsNA_cat <- traitsNA %>% 
  select(c('breeding_site', 'migration', 'vertical', 'diet', 'DielActivity'))

```


#subset trait data
Subset trait dataset according to the results from the regression analysis and save these.

```{r trait subset}
#generate final list with traits to use for further analysis and save the files as .csv
traits_final <- traits4 %>% 
  select(c('species', 'inter_litter_or_interbirth_interval_y', 'gestation_d', 'HabitatBreadth',
           'meanBodyMass', 'DielActivity', 'breeding_site', 'migration', 'vertical', 'diet')) %>% 
  rename(YearsInterBirth = 'inter_litter_or_interbirth_interval_y')
write.csv(traits_final, 'Traits/traits.csv')

#make a file for family and species
family <- traits4 %>% 
  select(c('species', 'family'))
write.csv(family, 'Traits/family.csv')

#make an iucn file
iucn <- traits4 %>% 
  select(c('species', 'IUCN'))
write.csv(iucn, 'Traits/iucn.csv')

```

##Trait imputation
Following method from Cooke et al. 2019 https://doi.org/10.1038/s41467-019-10284-z

#1. Upload trait data

```{r trait setup}
#set up trait data
trait <- read.csv('Traits/traits.csv')
#make categorical variables into numerical integers in order to be able to do imputation and functional analysis
unique(trait$DielActivity) #NCD, D, NC and CD
trait <- trait %>% 
  dplyr::select(-X) %>% 
  dplyr::mutate(migration = if_else(migration == 'marine', 1, 0)) %>% 
  dplyr::mutate(breeding_site = if_else(breeding_site == 'coastal', 0, 
                                        if_else(breeding_site == 'oceanic', 1, 2))) %>% 
  dplyr::mutate(vertical = if_else(vertical == 'benthic', 0,
                                   if_else(vertical == 'pelagic', 1, 2))) %>% 
  dplyr::mutate(diet = if_else(diet == 'plankton', 0, 
                               if_else(diet == 'inverts', 1,
                                       if_else(diet == 'inverts/fish', 2,
                                               if_else(diet == 'fish', 3, 4))))) %>% 
  dplyr::mutate(diet = as.integer(diet),
                migration = as.integer(migration)) %>% 
  dplyr::mutate(DielActivity = if_else(DielActivity == 'D', 1,
                                        if_else(DielActivity == 'CD', 2,
                                                if_else(DielActivity == 'NC', 3, 4))))
#check how many NAs are in the data - how many % per trait needs to be imputated
NAs_t <- colSums(is.na(trait))
NAs_gest <- NAs_t[3]/82*100 #22%
NAs_mig <- NAs_t[8]/82*100 #8.54%
#subset to the final 5 traits to check for how many species have NAs
tr5 <- trait %>% 
  select(-c('breeding_site', 'vertical', 'DielActivity', 'YearsInterBirth'))

NAs_species <- rowSums(is.na(tr5))
NAs_species <- as.data.frame(NAs_species)
NAs_species <- NAs_species %>% 
  filter(NAs_species != 0)
20/82*100 #24.39%

#upload family data
family <- read.csv('Traits/family.csv')
family <- family %>% 
  dplyr::select(-X)

#merge traits and family data
tr <- trait %>% 
  # add class data
  dplyr::left_join(family, by = "species")
```

#2. Upload phylogenetic data
phylogentic tree cetacean subset from MammalTree obtained from http://vertlife.org/phylosubsets/ (Upham et al 2019 'Inferring the mammal tree: Species-level sets of phylogenies for questions in ecology, evolution, and conservation.')

```{r phylogenetic tree}
# phylogenetic data
ceta_tree <- read.nexus('~/Sussex/thesis/Traits data/CetaceaTree/output.nex')
class(ceta_tree)
#computes the maximum clade credibility tree from a sample of trees
tree <- phangorn::maxCladeCred(ceta_tree)
PVR <- PVRdecomp(tree)
#select out species names in the order of the phylo trees
species <- PVR@phylo$tip.label
species <- as.data.frame(species)
#change species names
species$species <- sub("_", " ", species$species)
#make a dataframe with the phylogenetic eigenvectors and select the first 10 
eigen <- as.data.frame(PVR@Eigen$vectors)
eigen10 <- eigen %>% 
  select('c1':'c10')
#combine phylogenetic eigenvectors to species names and rename column names
pem <- cbind(species, eigen10)
oldnames <- c('c1', 'c2', 'c3', 'c4', 'c5', 'c6', 'c7', 'c8', 'c9', 'c10')
newnames <- c('V_1', 'V_2', 'V_3', 'V_4', 'V_5', 'V_6', 'V_7', 'V_8', 'V_9', 'V_10')
pem <- pem %>% 
  rename_at(vars(oldnames), ~ newnames)

#save phylogenetic eigenvectors dataframe
saveRDS(pem, 'Traits/pem_cetaceans.Rds')
#pem <- readRDS('Traits/pem_cetaceans.Rds')
```


```{r imputation}
#sort both data tables by binomial alphabetically
tr_sort <- tr[order(tr[,'species']), ]
pem_sort <- pem[order(pem[, 'species']),]
# join trait and phylogenetic data
tr_ceta <- dplyr::left_join(tr_sort, pem_sort, by = c('species'))

#change dataframe to get rid off errors - try making the variables more similar...
tr_ceta1 <- tr_ceta %>% 
  dplyr::mutate(meanBodyMass = meanBodyMass/1000000)

# missing data pattern
mice::md.pattern(dplyr::select(tr_ceta1, -species, -family, -c(V_1:V_10)))

# run multiple imputation
tr_mi_raw <- mice(tr_ceta1, m = 25, maxit = 100, seed = 20)
# method = pmm predictive mean matching

# summary of multiple imputation results including predictor matrix (which variables were used to predict missing values)
summary(tr_mi_raw)

# save: tr_mi_raw
saveRDS(tr_mi_raw, 'Traits/tr_mi_raw.rds')
# dataframe of raw multiple imputated trait data for mammals (data.frame)

tr_mi <- lapply(1:25, function(x) {
  out <- mice::complete(tr_mi_raw, action = x) %>% 
    dplyr::select(species:diet, -family) 
})

# save: tr_mi
saveRDS(tr_mi, 'Traits/tr_mi.rds')

```


