---
title: "Aquamaps"
author: "Jannika Oeke"
date: "6/28/2021"
output: 
  pdf_document:
    keep_tex: True

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Native Range maps for cetaceans
Maps downloaded from aquamaps.org for each cetacean species (n = 81) in .nc format and saved in a single folder.
Reference:
Kaschner, K. et al. (2019). AquaMaps: Predicted range maps for aquatic species. Retrieved from https://www.aquamaps.org.

```{r libraries, message=FALSE, warning=FALSE}
library(ncdf4)
library(raster)
library(dplyr)
library(tidyverse)
library(RColorBrewer)
library(maptools)
library(myplot)
```

# Input nc files
Input all nc files and save as a raster stack

```{r data input nc, eval=FALSE, include=TRUE}
#specify path to the .nc files
ncpath <- '~/Sussex/thesis/Aquamaps/nc'
nc_files <- list.files(path = ncpath, pattern = '.nc$', full.names = TRUE) 

NC <- nc_open(nc_files)
print(NC)

# Get whale names
whale_names <- vector()
for (i in 1:length(nc_files)){
  whale_names[i] <- str_extract(nc_files[i], "(?<=nc/)[^;]*(?=.nc)")
}

#set extent for raster stack
e <- extent(-180, 180, -90, 90)
#make empty list for rasters
rasters <- list()
#extract each raster and extend or crop extent accordingly
for(i in 1:length(NC$filename)){
  rasters[[i]] <- raster(NC$filename[[i]]) %>%
    extend(., e) %>%
    crop(., e)
}

#close NC files
nc_close(NC)

#stack all individual rasters
whales <- raster::stack(rasters)

names(whales)
#rename individual layers from probability.1 etc to species names using the vector created earlier
names(whales) <- whale_names

#plot individual maps 
plot(whales[[2]])

#save raster stack
saveRDS(whales, 'RES_rasterstack.Rds')
#whales <- readRDS('RES_rasterstack.Rds')

#set colour and breaks for pretty plot
col <- brewer.pal(6, 'GnBu')
brk <- c(0, 0.01, 0.2, 0.4, 0.6, 0.8, 1)
#plot map for a single species
p1 <- myplot({
  plot(whales[[41]], #humpback
     breaks = brk, col = col,
     main = 'Megaptera novaeangliae Native Range', 
     xlab = 'Longitude', ylab = 'Latitude',
     ylim = c(-90,90),
     legend.args = list(text = 'RES', font = 2, cex =0.8))
})

plotsave(p1, file = 'RES_humpback.png', width = 15, height = 10, units = 'cm')
```

#Convert RES to likely presence/absence
Relative environmental suitability (RES) (= probability of occurrence) ranges from 0-1, define >=0.6 as (likely) present = 1, rest as (likely) absent = 0

```{r presence, eval=FALSE, include=TRUE}
#change RES probability to presence
#make a matrix with the conditions to reclassify the data
#3 column table: 1 column is the 'from' 2 column 'to' and 3 column 'becomes'
#here: all values >0 and <= 0.59 become 0 AND >0.59 and <= 1 become 1

fct <- c(0, 0.59, 0, 0.59, 1, 1)
fct1 <- matrix(fct, ncol = 3, byrow = T)

#create empty list to create new rasters into
presence <- list()
#loop reclassifying the data in each raster layer 
#1. changing all NA to 0
#2. change RES to presence/absence
for(i in 1:length(whales@layers)){
  presence[[i]] <- reclassify(whales[[i]], cbind(NA, 0)) %>%
  reclassify(., fct1) 
}
#stack new raster layers
whales_pres <- stack(presence)
#plot a few individual layers
plot(whales_pres[[2:5]])
#save raster stack 
saveRDS(whales_pres, 'pres_rasterstack.Rds')
#whales_pres <- readRDS('pres_rasterstack.Rds')

#plot single layer
plot(whales_pres[[41]])

p2 <- myplot({
  plot(whales_pres[[41]], #humpback
     main = 'Megaptera novaeangliae Presence (RES \u2265 0.6)', #RES bigger equal 0.6 
     xlab = 'Longitude', ylab = 'Latitude',
     ylim = c(-90,90),
     legend = FALSE)
})
plotsave(p2, file = 'pres_humpback.png', width = 15, height = 10, units = 'cm')

#sum all presences together
sum_pres <- stackApply(whales_pres, indices = 1, fun = sum)
#save summary
saveRDS(sum_pres, 'sum_pres_rasterstack.Rds')
#sum_pres <- readRDS('sum_pres_rasterstack.Rds')
sum_pres_df <- sum_pres@data
plot(sum_pres)
sum_pres@data@max #41

col2 <- brewer.pal(8, 'YlGnBu')
col3 <- viridis::viridis(11)
brk2 <- c(0, 1, 5, 10, 15, 20, 25, 30, 35, 40, 41)
#load world map from maptools package
data('wrld_simpl')

p3 <- myplot({
  plot(sum_pres, breaks = brk2, col = col3,
     #main = 'Cetacean Species Richness', 
     xlab = 'Longitude', ylab = 'Latitude',
     ylim = c(-90,90),
     legend.args = list(text = 'Number of Species', font = 2, line = 0.5, side = 2, cex =0.8))
  #add continents in white without black borders of countries
  plot(wrld_simpl, 
       col = 'white',
       border = 'white', 
       add = TRUE)
})

plotsave(p3, file = 'sum_pres-wrld3.png', width = 15, height = 10, units = 'cm')
```
